---
title: "Get started"
author: "Mauro Lepore"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
set.seed(1014)
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  # Figures (http://r4ds.had.co.nz/)
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,
  fig.show = "hold"
)
```

## Overview

__fgeo.data__ provides open datasets of ForestGEO. To get started run this:

```{r}
# install.packages("remotes")
# remotes::install_github("forestgeo/fgeo.data")
library(fgeo.data)
```

(See [Installation](https://forestgeo.github.io/fgeo.data/index.html#installation))

For an index of all available datasets and access see [Reference](https://forestgeo.github.io/fgeo.data/reference/index.html).

## Data dictionary

Find descriptions searching by table or column name.

```{r, echo=FALSE}
DT::datatable(data_dictionary)
```

## Example

### ViewFullTable (`vft`)

```{r}
str(luquillo_vft_4quad, give.attr = FALSE)
```

### ViewTaxonomy (`taxa`)

```{r}
str(luquillo_taxa, give.attr = FALSE)
```

### Census (`tree` and `stem`)

```{r}
str(luquillo_tree6_random, give.attr = FALSE)
```

`luquillo_stem_1ha` combines multiple censuses in a sigle dataframe.

```{r}
lapply(luquillo_stem_1ha, function(column) head(unique(column)))
```
This dataset allows summaries by census. For example, you may want to know which quadrats have the largest number of alive trees per census:

```{r}
library(ggplot2)
library(dplyr)

smry <- luquillo_stem_1ha %>% 
  filter(status == "A", !is.na(dbh)) %>% 
  group_by(CensusID, quadrat) %>% 
  summarise(n = n())

top_n(smry, 1)
```

```{r}
ggplot(top_n(smry, 1), aes(CensusID, n)) + 
  geom_col(fill = "lightgreen") +
  geom_text(aes(label = quadrat)) +
  labs(title = "Quadrat with greatest number of alive trees per census")
```

### Elevation (`elevation`)

```{r}
str(luquillo_elevation)

head(luquillo_elevation$col)
```

## Reading data from .csv fiels (e.g. the database output of ViewFullTable)

To safely read raw data from .csv files, you need to know about the most important arguments of the functions you use to read the data. Below is a small example but you may also want to read these articles, showing two ways to read data:

* With [__readr__](https://fgeo.netlify.com/2018/03/13/2018-03-13-import-dataset/) (recommended).
* [With only base R](https://fgeo.netlify.com/2018/03/15/2018-03-15-how-to-read-data-safely-with-only-base-r/).

```{r}
# ForestGEO's database delivers tab delimited files with extension .csv.
# Instead of an url, you likely have a path to a directory in your computer.
url <- "http://bit.ly/fgeo-data-luquillo-taxa"

# TO AVOID ERRORS
# Specify columns class explicitely (for an example showing only some columns)
column_classes <- c(
  ViewID = "integer", 
  SpeciesID = "integer", 
  SubspeciesID = "integer", 
  Family = "character", 
  Mnemonic = "character"
)
# Specify how to interprate missing values
na_strings <- c("", "NA", "NULL")

raw_luquillo_taxa <- read.delim(
  file = url, 
  na.strings = na_strings, 
  colClasses = column_classes
)

some_rows <- some_columns <- 1:5
raw_luquillo_taxa[some_rows, some_columns]

# Note these are the same as `column_classes` defined above
sapply(raw_luquillo_taxa[some_rows, some_columns], class)
```
